OneD: increasing reproducibility of Hi-C Samples with abnormal karyotypes

### 2017-06-01 main.tex
Wrote the first few paragraphs of the introduction.

High throughput chromosomal conformation capture (HiC) experiments allow us to explore the spatial organization of chromatin in cells \citep{lieberman2009comprehensive}. Contrary to other 3C techniques, HiC is unbiased in the sense that all possible pairwise interactions between restriction fragments might be interrogated. However, this does not guarantee absence of bias. For instance, genome features as CG content, restriction enzyme site distribution and sequence uniqueness (related to mappability) have been described to modify the result in such experiments \citep{yaffe2011probabilistic}.

Several strategies have been proposed to remove biases in HiC experiments \citep{schmitt2016genome}, either modeling explicitly an a-priori set of known genomic features as potential sources of bias \citep{yaffe2011probabilistic,hu2012hicnorm} or using algorithms to implicitly correct for unknown biases \citep{imakaev2012iterative}. Notably, the latter has a broader usage due to its fast implementation. Both types of methods are suitable for normal cells but they were not designed for cells with abnormal karyotypes. Although there have been some proposals that consider aneuploidy \citep{wu2016computational}, to our knowledge, there is a lack of methodology dealing with aberrant karyotypes (common in cancer cell lines) that present abnormalities more complex than entire chromosome duplications.

>>>>

One of the crown achievements of modern biology was to realize that
genomes have an underlying three-dimensional structure contributing to
their activity (CITE). In mammals, this organization plays a key role in
guiding enhancer-promoter contacts (CITE), in V(D)J recombination (CITE)
and in X chromosome inactivation (CITE). One of the most significant
breakthroughs towards this insight was the development of the high
throughput chromosomal conformation capture (Hi-C) technology, assaying
chromosomal contacts at a genome-wide scale
\citep{lieberman2009comprehensive}. Exploring the spatial organization of
chromatin inside the cell is now common and Hi-C is part of the standard
molecular biology toolbox (CITE).

Contrary to the precursor technologies 3C, 4C and 5C \citep{de2012decade},
Hi-C interrogates all possible pairwise interactions between restriction
fragments. However, this does not guarantee the absence of bias. For
instance, local genome features such as the G+C content, the availability
of restriction enzyme sites and the mappability of the sequencing reads
have been shown to modify the results \citep{yaffe2011probabilistic}. In
addition, the general experimental biases such as batch effects and
protocol variation also apply. It is thus important to normalize
Hi-C data in order to remove biases and artifacts, so that they are not
confused with biological signal.

Several methods have been proposed to remove biases in Hi-C experiments
\citep{schmitt2016genome}. The first strategy is to model biases
explicitly from a defined set of local genomic features, such as the G+C
content and the availability of restriction sites. This approach is used
in the method of \cite{yaffe2011probabilistic} and in Hicnorm by
\cite{hu2012hicnorm}. The second strategy is to implicitly correct for
unknown biases by enforcing some regularity on the data. This approach is
used in the ICE method of \cite{imakaev2012iterative}, where the total
amount of contacts is imposed to be equal for every bin. ICE is currently
the most popular method, which is in part due to the speed of the
available implementation.

Both strategies are suitable for cells with a normal karyotype, but they
were not designed for cell types with karyotypic aberrations.

Although there have been some proposals that consider aneuploidy
\citep{wu2016computational}, to our knowledge, there is a lack of
methodology dealing with aberrant karyotypes (common in cancer cell lines)
that present abnormalities more complex than entire chromosome
duplications.
